---
title: "Course project 1"
author: "zengyujia"
date: "2025-03-21"
output: html_document
---

## Introduction 

This is the peer-graded assignment for the Reproducible Research Course by Johns Hopkins University at Coursera, which is the 5th out of 10 courses in the [Data Science Specialization](https://www.coursera.org/specializations/jhu-data-science). 

The objective of this assignment is to make a reproducible report of an individual's activity over a two month period, measured in number of steps. The report will be generated using [knitr](https://yihui.name/knitr/). 

As provided in Professor Peng's [original repository](https://github.com/rdpeng/RepData_PeerAssessment1), this repository is self-contained — all data and assignment instructions necessary to reproduce this work are available in the same place. 

## Setting up the report
First, we need to setup the report options using, and install —if necessary— and load all the required packages to successfully complete this assignment. 

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
## Loading the necessary package for reproducing the assingment
if (!require(knitr)) {
    install.packages("knitr")
}
library(knitr)

if (!require(tidyverse)) {
    install.packages("tidyverse")
}
library(tidyverse)

## Setting up all code chunks according to the assignment specs
knitr::opts_chunk$set(
    eval = TRUE,
    echo = TRUE,
    tidy = TRUE,
    results = "markup",
    include = TRUE,
    message = FALSE,
    warning = FALSE,
    knitr.table.format = "markdown", 
    tidy.opts = list(width.cutoff = 80), 
    fig.align = "center", 
    fig.path = "figure/", 
    highlight = TRUE
)
```

## Loading and preprocessing the data 

Instructions: 
    Show any code that is needed to
    1. Load the data (i.e. read.csv())
    2. Process/transform the data (if necessary) into a format suitable for your analysis

In the original repository, the data set is compressed in a ZIP file, which we unzip —if necessary— and load the resulting CSV file into the working environment. 

No pre-processing was necessary for this data set, as it is already tidy. 

```{r unzip-load-preprocess data}
## If necessary, unzipping the data file, and loading it
if (!file.exists("activity.csv") &
    file.exists("activity.zip")) {
    unzip("activity.zip")
    activity.data <- read_csv(file = "activity.csv")
} else if (file.exists("activity.csv")) {
    activity.data <- read_csv(file = "activity.csv")
} else {
    message("Activity Monitoring Data (default) from Rep Research course was not found")
}
data <- read.csv('activity.csv')
sum_steps <- data %>%
  group_by(date) %>%
  summarise(steps = sum(steps))

mean_median_steps <- sum_steps %>%
  summarise(Mean = mean(steps, na.rm = TRUE),
            Median = median(steps, na.rm = TRUE)) %>%
  gather()
ggplot(data = sum_steps, 
       mapping = aes(x = steps, 
                     y = ..density..)) +
    geom_histogram(aes(weight = steps),
                   fill = "steelblue",
                   colour = "white",
                   alpha = 0.8) +
    geom_density(fill = "steelblue",
                 colour = NA,
                 alpha = 0.2) +
    ggtitle(list(
    title = "Total Number of Steps",
    subtitle = "Distribution of Total Number of Steps Taken Each Day",
                 x = "Steps", 
                 y = "Density")) +
    geom_vline(data = mean_median_steps, 
               aes(xintercept = value, linetype = key), 
               size = 0.5,
               alpha = 0.8,
               colour = c("salmon", "royalblue")) +
    scale_linetype_discrete(name = "Number of Steps") +
    theme_bw() 
```
## Mean and Median
```{r descripitive statistics}
sum_steps %>%
    summarise(
    "Mean number of steps taken per day" = round(mean(steps, na.rm = TRUE)),
    "Median number of steps taken per day" = round(median(steps, na.rm = TRUE))
    ) %>%
    kable(align = "c")
```
```{r activity pattern}
avg_steps_day <- data %>%
  group_by(interval) %>% 
  summarise(steps = round(mean(steps, na.rm = TRUE))) %>%
  filter(!is.nan(steps))
max.steps <- avg_steps_day[which.max(avg_steps_day$steps),]
ggplot(data = avg_steps_day, mapping = aes(x = interval, y = steps)) +
    geom_line(size = 0.5,
    colour = "steelblue",
    alpha = 0.9) +
        ggtitle(list(title = "Daily Activity Pattern", 
                     subtitle = "Average number of steps taken by 5-min interval, across all days",
                     y = "Number of Steps", 
                     x = "Interval")) +
        geom_point(data = max.steps, 
                   mapping = aes(x = interval, y = steps), 
                   size = 4, 
                   alpha = 0.5, 
                   colour = "salmon") +
    theme_bw()

```

## NA counts
```{r na_count}
nas.df <- data %>%
  sapply(is.na) %>%
  as_tibble() %>%
  summarise(Steps = sum(steps),
            Date = sum(date),
            Interal = sum(interval))
row.names(nas.df) <- c('Numbers of NAs')
nas.df %>% 
    kable(caption = "Number of Missing Values of each Variable",
          align = "c")

```


## NA imputation
```{r na imputation}
non.NA <- data %>%
  group_by(interval) %>% 
  summarise(steps = round(mean(steps, na.rm = TRUE, trim = 0.05)))
imp.activity.data <- full_join(x = data, y = non.NA, by = 'interval') %>%
  mutate(steps = ifelse(test = is.na(steps.x), yes =  steps.y, no = steps.x)) %>%
  select(steps, date, interval)
```

## Hist of compelete data
```{r histogram of complete steps}
steps.day.df <- imp.activity.data %>%
    group_by(date) %>%
    summarise(steps = sum(steps))
mean.median.steps <- steps.day.df %>%
    summarise(
        Mean = round(mean(steps, na.rm = TRUE)),
    Median = round(median(steps, na.rm = TRUE))
    ) %>% 
    gather()
    
ggplot(data = steps.day.df, 
       mapping = aes(x = steps, 
                     y = ..density..)) +
    geom_histogram(aes(weight = steps),
                   fill = "steelblue",
                   colour = "white",
                   alpha = 0.8) +
    geom_density(fill = "steelblue",
                 colour = NA,
                 alpha = 0.2) +
    ggtitle(list(
    title = "Total Number of Steps",
    subtitle = "Distribution of Total Number of Steps Taken Each Day",
                 x = "Steps", 
                 y = "Density")) +
    geom_vline(data = mean.median.steps, 
               aes(xintercept = value, linetype = key), 
               size = 0.5,
               alpha = 0.8,
               colour = c("yellow", "red")) +
    scale_linetype_discrete(name = "Number of Steps") +
    theme_bw() 
```
  